---
title: | 
  | **The R *iconr* package** 
  | a tutorial
bibliography: references.bib
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r, echo=FALSE}
# print(getwd())
# print(file.path("../logo", "iconr_logo.png"))
htmltools::img(src = knitr::image_uri(file.path("img", "iconr_logo.png")), 
# htmltools::img(src = knitr::image_uri(file.path(R.home("doc"), "html", "logo.jpg")), 
               alt = 'logo', 
               style = 'position:absolute; top:0; right:0; padding:10px;')
```

---

<style>
.figure {
   margin-top: 0px;
   margin-bottom: 40px;
}
table {
    margin-top: 0px;
    margin-bottom: 24px;
}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(kableExtra)
library(dplyr)

solanas.vor.path <- "img/all_process.gif"
```
  

 the ***iconr*** embedded training dataset, others datasets are available. To facilitate the starting, ready to use QGIS projects (.qgz) are also provided for each decoration:
```{r, echo = F}

library(httr)
req <- GET("https://api.github.com/zoometh/iconr/docs/datasets/git/trees/master?recursive=1")
stop_for_status(req)
filelist <- unlist(lapply(content(req)$tree, "[", "path"), use.names = F)
grep("Matteo/literature/", filelist, value = TRUE, fixed = TRUE)

url.root <- "https://raw.github.com/zoometh/iconr/master/docs/datasets/"
doss1 <- paste0(url.root, "Valcamonica/")
file1 <- paste0(doss1, "edges.tsv")
download.file(url = URL, destfile=basename(URL))
gh.root <- "https://downgit.github.io/#/home?url=https://github.com/zoometh/iconr/tree/master/docs/datasets/"
url.df <- c(paste0(gh.root, "stele%20bas%20aragon"),
            paste0(gh.root, "Valcamonica"))
url.ref <- c("https://doi.org/10.3989/gladius.2013.0002",
             "https://scholar.google.fr/scholar?hl=fr&as_sdt=0%2C5&q=Le+Pietre+degli+Dei%2C+menhir+e+stele+dell%27Eta+del+Rame+in+Valcamonica+e+Valtellina&btnG=")
df.datasets <- data.frame("download" = c("stele bas aragon", "Valcamonica"),
                          "name" = c("Estelas Ibericas con Lanzas", "Ossimo"),
                          "nb.decor" = c("5","3"),
                          "sort.description" = c("Late Iron Age stelae with ranges of spears, writings, etc.",
                                                 "Chalcolithic rock-art of Ossimo, Valcamonica"),
                          "drawings.references" = c("Vargas 2013","Fedele 1994"),
                          stringsAsFactors = F)
# row.names(df.datasets) <- df.datasets$family
df.datasets %>% 
  mutate(download = cell_spec(download, "html", link = url.df)) %>%
  mutate(drawings.references = cell_spec(drawings.references, "html", link = url.ref)) %>%
  kable("html", escape = FALSE, row.names = F) %>%
  kable_styling(bootstrap_options = c("hover", "condensed"), full_width = F)
```


## Start to work with the ***incor*** package

Now we have at least one decoration modeled with a geometric graph, we can start to use the ***incor*** package. Its latest development version with its documentation (vignette) can be downloaded from GitHub

```{r down,eval=FALSE, echo=TRUE}
devtools::install_github("zoometh/iconr", build_vignettes=TRUE)
```

And load the package

```{r echo=TRUE}
library(iconr)
```

To start using the package, you have first to locate your working directory. For example:

```{r echo=TRUE}
dataDir <- paste0(getwd(), "/extdata")
```

The you can start with the function `plot_dec_grph()` and specifying the extensions of the nodes and edges data (`shp`)

```{r echo=TRUE, out.width="60%", fig.width=8, fig.asp=750/666, fig.align="center", warning=FALSE, fig.cap="Abela stelae"}
# Decoration to be plotted
site <- "Abela"
decor <- "Abela"
# Read nodes, edges, and decorations
nds.df <- read_nds(site, decor, dataDir, format = "shp")
eds.df <- read_eds(site, decor, dataDir, format = "shp")
imgs <- read.table(paste0(dataDir, "/imgs.tsv"),
                   sep="\t", stringsAsFactors = FALSE, header = T)

# Save the plot of nodes and edges with node variable "type" as labels
# in png image format and return the image file name.
plot_dec_grph(nds.df, eds.df, imgs,
              site, decor,
              dir = dataDir,
              nd.var = "type")
```

