---
title: "Graph and Spatial Analysis of Iconography for Archaeology"
author: "Thomas Huet"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Graph and Spatial Analysis of Iconography for Archaeology}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
references:
- id: Huet18
  title: Geometric graphs to study ceramic decoration
  author:
  - family: Huet
    given: Thomas
  container-title: Proceedings of the 44th Conference on Computer Applications and Quantitative Methods in Archaeology, CAA 2016
  publisher: Archaeopress
  page: 311-324
  type: article-journal
  URL: 'https://www.researchgate.net/publication/343514122_Geometric_graphs_to_study_ceramic_decoration'
  issued:
    year: 2018
---

```{r, include = FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.pos = 'H'
)
```

# Introduction

`decorr` package is used to study ancient iconography with spatial and network analysis indexes [@Huet18].

A relative neighborhood graphs (RNG) is constructed over the decoration image to store spatial relationships between graphical units (GUs). To compare pairwises of decorations (ie, RNG) the `decorr` package has function to count and plot common nodes and/or common edges.

# Install

Install the `decorr` package and load it

```{r load_decorr}
# devtools::install_github("zoometh/iconr")
library(decorr)
```

## Contents

**functions**

```{r ls_decorr}
print(ls("package:decorr"))
```

**external data**

```{r ushape.profile, fig.align="center",fig.cap="\\label{fig:figs}U-shape profile"}
doss <- system.file("extdata", package = "decorr")
print(list.files(doss))
```

... you may change `doss` to the path of your working directory

* 'drawing': load an image with the `magick` package

```{r drawing, out.width = "100%", fig.align="center"}
drawing <- magick::image_read(system.file("extdata",
                                          "Cerro_Muriano.Cerro_Muriano_1.jpg",
                                          package = "decorr"))
drawing <- magick::image_trim(drawing)
plot(drawing)
```

* 'nodes': a dataframe or a shapefile of the nodes

```{r nodes.df}
doss <- system.file("extdata", package = "decorr")
site <- "Cerro Muriano"
decor <- "Cerro Muriano 1"
nds.df <- read_nds(doss = doss, site = site, decor = decor, nodes = "nodes", dev = ".tsv") 
kable(nds.df)
```

The `read_nds()` function allows to read `.tsv`, `.csv` (with ';' as values separator) or `.shp` files.

  + column names  
  
    - `id`: identifiers of nodes (a unique number) 
    - `type`: type of nodes    
    - `x, y`: coordinates of nodes, these fields are only needed for the `.tsv`, `.csv` files    

* 'edges': a dataframe or a shapefile of the nodes  

```{r edges.df}
doss <- system.file("extdata", package = "decorr")
site <- "Cerro Muriano"
decor <- "Cerro Muriano 1"
eds.df <- read_eds(doss = doss, site = site, decor = decor, edges = "edges", dev = ".csv")
kable(eds.df)
```
  
  + column names 
  
    - `id`: identifiers of edges (a unique number) 
    - `type`: type of edges (`=`: normal edge, `+`: node `b` is an attribute of node `a`, `>`: node `a` overlap node `b`)  
    - `a`: *id* of the first node  
    - `b`: *id* of the second node  
    - `xa, ya`: coordinates of the first node  (only needed for the `.tsv`, `.csv` files)  
    - `xb, yb`: coordinates of the second node (only needed for the `.tsv`, `.csv` files)  
    

## Example

Load nodes, edges and create graph for "type" field. This field stores intrinsic values of the GU. It could be its type, its color, its technique, and so on.

```{r read_img}
site <- "Cerro Muriano"
decor <- "Cerro Muriano 1"
nds.df <- read_nds(doss = doss,site = site,decor = decor,nodes = "nodes",dev = ".shp")
eds.df <- read_eds(doss = doss,site = site,decor = decor,edges = "edges",dev = ".shp")
img.graph <- plot_dec_grph(nds.df = nds.df,
                           eds.df = eds.df,
                           doss = doss,
                           site = "Cerro Muriano",
                           decor = "Cerro Muriano 1",
                           lbl.txt = "type",
                           shw = c("nodes","edges"))
```

`img.graph` is a `r class(img.graph)`  

Plot the RNG graph showing the "type" of the nodes

```{r img.graph.plot, out.width = "100%",fig.align="center"}
img.graph <- magick::image_trim(img.graph)
plot(img.graph)
```

## References

